### 2016-04-08 paper/paper.tex
alg

%------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%------------------------------------------------

\documentclass[twoside]{article}

\usepackage{graphicx}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{listings}
\lstset{
breaklines=true,
tabsize=2
}

\usepackage{amsmath,amssymb,amsthm} % Mathematical Symbols, styles, etc

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\usepackage[sc]{mathpazo} % Use the Palatino font
% Output encoding
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
% Input encoding
\usepackage[utf8]{inputenc} % UTF-8 character encoding stuff
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{R.L.H. Fontein, H.D. van Wieren, T. Kerkhoven: \shorttitle} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

% Bibliography
\usepackage[backend=bibtex, sorting=none]{biblatex}
\bibliography{references.bib}

% Appendices
\usepackage[toc,page]{appendix} % appendix

% Additional column type
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

%------------------------------------------------
%	TITLE SECTION
%------------------------------------------------

\newcommand{\articletitle}{Comparison between Time Complexity of Different Color Refinement Algorithms in Graph Isomorphism}
\newcommand{\shorttitle}{Color Refinement Time Complexity in Graph Isomorphism}

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{\articletitle}} % Article title

\author{
\large
\textsc{Rick Fontein, Huub van Wieren, and Tim Kerkhoven}\\[2mm] % Your name
\normalsize University of Twente \\ % Your institution
\normalsize \href{mailto:r.l.h.fontein@student.utwente.nl}{r.l.h.fontein@student.utwente.nl},
\href{mailto:h.d.vanwieren@student.utwente.nl}{h.d.vanwieren@student.utwente.nl},\\
\normalsize \href{mailto:t.kerkhoven@student.utwente.nl}{t.kerkhoven@student.utwente.nl}% Your email addresses
}

\date{\today}

%------------------------------------------------

\begin{document}

\thispagestyle{empty}
\maketitle % Insert title

%------------------------------------------------
%	ABSTRACT
%------------------------------------------------

\begin{abstract}

\noindent The worst-case asymptotic time complexity for a basic and fast color refinement algorithm are calculated to be $O(N^3)$ and $O(M\cdot{}log(N)$ respectively. Computational results for a three paths type graph indicate different time complexities, at least for that specific type of graph. It can be concluded that either the algorithms are faster for the three paths graphs, or the calculations for the time complexity are incorrect.

\end{abstract}

%------------------------------------------------
%	ARTICLE CONTENTS
%------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

%------------------------------------------------
\section{Introduction}
\label{intro}
This paper will compare the time complexity of two color refinement algorithms and confirm them with computational results. The first algorithm, henceforth referred to as the basic algorithm, uses the color refinement algorithm as shown by J. de Jong\cite{presentation:slidesPartI}, while the other algorithm, henceforth referred to as the fast algorithm, is based on DFA minimization as shown by P. Bonsma\cite{presentation:slidesPartIII}. The algorithms have been implemented and their computation times on test sets will be used to confirm the time complexity.

Color refinement algorithms are used in solving the graph isomorphism problem, which is the computational problem of determining whether two finite graphs are isomorphic. It is one of a very small number of problems belonging to NP neither known to be solvable in polynomial time nor NP-complete: it is one of 12 such problems listed by Garey \& Johnson (1979)\cite{book:gareyJohnson1979}, and the only one of that list whose complexity remains unresolved.\cite{website:wikiGI}

%------------------------------------------------
\section{Scope \& Objectives}
\label{scope}
The aim of this research paper is to calculate and compare the worst-case asymptotic time complexity of both color refinement algorithms and verifying them using computational results. These results will be used to examine the differences in the growth of the computation time for a specific type of graph.

This paper will describe both algorithms, give the calculations for the asymptotic worst-case time complexity, as well as an examination of the computational results.
%------------------------------------------------
\section{Methods}
\label{methods}
The time complexity of both algorithms was calculated by hand. To verify the results, a computational experiment was performed. In this experiment, both algorithms were tested with various sizes of a graph consisting of three paths. The paths are connected to each other on both of their ends. In the experiment the only thing changing is the length of these paths.

The testing was all done on the same machine, under the same circumstances. Every test was performed three times in order to prevent outliers caused by external events.

\subsection{Testing}
\label{methods:testing}
In the experiment, both algorithms were tested with several sizes of the same type of graph, as described above. Each algorithm was timed three times. The lowest value of the three was selected as the most accurate, as the error margin for timing algorithms on a system is one-sided: it can only go slower than its optimal value.

The python library used for timing the algorithm is \emph{timeit}. The manner in which it was used is the following:
\begin{lstlisting}
	min(Timer(algorithm_instance).repeat(3, 1))
\end{lstlisting}

Only the algorithms themselves and the initial coloring were timed, initialization and other parts of the code were not included.

The fast color refinement algorithm used for this paper is made by W. van Veelen, N. Goldsmid, E. Khan, and E. Koster \cite{misc:group4}, because the original implementation had some issues.

%------------------------------------------------
\section{Algorithms}
\label{algs}
%% TODO %%
% Describe algorithms, show concise calculations of time complexity
%%
In the following subsection, a general overview and an analysis of the algorithms is given:
The computation time of the algorithms increases for iterations over the vertices in a graph. To calculate the worst-case time complexity of the algorithms, the number and places of the iterations over the vertices are analyzed.

The input and output for the algorithms are the same. The input is a graph with a possibly unstable  coloring. The output is a graph with stable coloring. However, the graphs do have a different structure to support their respective algorithm's refinement operation. We do not discuss the initialisation of these structures.

A coloring $\alpha$ of a graph $G$ is called stable if for all $u, v \in V(G)$ it holds that: if $\alpha(u) = \alpha(v)$, then $u$ and $v$ have identically colored neighborhoods.

\subsection{Basic}
The basic algorithm gives all vertices in a graph with the same color but differently colored neighborhoods a separate color. This is done iteratively until the graph has a stable coloring. The determining factor of this method is that all vertices are checked each iteration.

The basic implementation has several places where iterations over vertices take place. The code contains one main loop that has a worst-case of N iterations. Thereby the time complexity so far is $O(\boldsymbol{N})$, where $N = V(G)$. In this main loop an iteration takes place over all the vertices in the graph. The time complexity is now $O(N \cdot \boldsymbol{N})$. Each vertex is checked to see if it needs an update. If so, a new color dictionary is made and a new loop over the dictionary takes place. This makes the time complexity is $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary. Every color class in the dictionary is checked to see if the configuration has already been assigned a new color. If the the configuration is not yet know, a new color is assigned. In worst-case this new color operation costs N. The time complexity is $O((N \cdot N)(N+N+\boldsymbol{N}))$ so far.

After the first loop over the vertices in the main loop, the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary. The time complexity worst-case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$. At the end, the worst-case time complexity is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.

\subsection{Fast}
With fast color refinement the colorsclasses that could be used to refinement other color classes are tracked by putting them in a queue. If a color class is being refined, the class is split into two different classes, with one class keeping the old color. The algorithm terminates with a stable coloring if no color class needs refinement, in other words the queue is empty.

Fast color refinement is also analyzed. $N$ is again defined as $N=|V(G)|$. Determined is the number of times a vertex $v$ is refined in its worst-case. The color class of $v$ splits at least in half if it is getting refined. The split class with the most vertices in it keeps the original color. If vertex $v$ stays in this original class, $v$ is not refined by definition. It follows that if $v$ gets refined, the degree of the previous color class is at most two times the degree of its new class. Let $k$ be the times $v$ is refined in the algorithm, then $2^k$ is the degree of $v$'s original color class.  The degree of the color class is worst-case $N$ if all vertices have the same amount of edges, so worst-case $k=log_2 N$. As shown by P. Bonsma in one of his lectures. \cite{presentation:slidesPartIII}

With the number of edges in a graph times the number of vertices in that graph, the time complexity can be defined.\cite{presentation:slidesPartIII} The time complexity of fast color refinement is $O(m\cdot log n)$.

%------------------------------------------------
\section{Results}
\label{results}
%% WIP %%
% Give and describe results
%%
% Table for results can be found at appendices
The measured computation times are visualized in \autoref{fig:basiconly} and \autoref{fig:fastonly}, the actual times can be found in \autoref{table:measurements}.

\subsection{Basic Algorithm}
\autoref{fig:basiconly} shows the measured computation times of the basic color refinement algorithm for the three paths graphs with path lengths 5 through 5120. The larger sizes of this graph would have taken too long to compute within reasonable time.

\subsection{Fast Algorithm}
\autoref{fig:fastonly} shows the measured computation times of the fast color refinement algorithm for the three paths graphs with path lengths 5 through 20480. Larger variants of the graph were not available and therefore not used.

%------------------------------------------------
\section{Discussion} %Change section name?
\label{disc}
%% WIP %%
% Describe the effectiveness of the algorithms on various types of graphs, and examine the differences
%%
To verify the results of the time complexity calculations, the measurements have to be compared to functions of the expected time complexity. To more clearly see the relation between the two, diagrams with the relative size of the two were created.

\subsection{Basic Algorithm}
As the calculated worst-case time complexity of the basic color refinement algorithm is $O(N^3)$, \autoref{fig:basicfunctions} shows the measurements and the function $f(n)=n^3$. \autoref{fig:basicgrowth} then plots the relative size of the two on a logarithmic scale. It shows that $f(n)=n^3$ grows significantly faster than the measurements. Looking for a better fit, $f(n)=n^2$ was added to both diagrams. \autoref{fig:basicgrowth} indicates that the relative size is slowly decreasing and the growth of the measurements is approaching the growth of $f(n)=n^2$.

\subsection{Fast Algorithm}
As the calculated worst-case time complexity of the fast color refinement algorithm is $O(M\cdot{}logN)$, \autoref{fig:fastfunctions} shows the measurements and the function $f(n)=n\cdot{}log(n)$. \autoref{fig:fastgrowth} then plots the relative size of the two on a logarithmic scale. It shows that $f(n)=n\cdot{}log(n)$ grows faster than the measurements. Looking for a better fit, $f(n)=n$ was added to both diagrams. \autoref{fig:fastgrowth} indicates that $f(n)=n$ and the measurements are growing at the same rate, which implies that the time complexity for this specific type of graph might be closer to $\Theta(N)$.

%------------------------------------------------
\section{Conclusions}
\label{concl}
The most obvious conclusion that can be drawn from the worst-case time complexity and computational results is that fast refinement is a faster algorithm than basic refinement. However, the computational results are compared with their calculated worst-case complexity some differences were observed.

The determined time complexity of basic refinement is $O(N^3)$. When looking at the relative growth, it seems that the algorithm approaches $O(N^2)$. There could be two causes:
\begin{compactitem}
	\item The determined complexity of $O(N^3)$ was faulty.
	\item The structure of the three paths graphs causes the basic algorithm to approach a $O(N^2)$ time complexity. $O(N^3)$ was determined as worst-case.
\end{compactitem}

The determined time complexity of the fast algorithm is $O(MlogN)$. When looking at the relative growth, it seems that the algorithm approaches a linear complexity. This can be attributed to the structure of the type of graphs used, three paths. $O(MlogN)$ was determined as worst-case.

%------------------------------------------------
%	REFERENCE LIST
%------------------------------------------------
\label{references}
\printbibliography
%bibliography{references}

%------------------------------------------------
%	APPENDICES
%------------------------------------------------
%% TODO %%
% Put all the things that should be in the paper, but don't fit the main part here
%%

% Everything EXCEPT multicols can be removed from the appendices if none are added.
\end{multicols}

\begin{appendices}
\label{app}

%------------------------------------------------
\section{Algorithms}
\label{app:algs}
\begin{algorithm}
    \begin{algorithmic}[1] % The number tells where the line numbering should start
        \Procedure{BasicColorRefinement}{$G$}
            \While{not stable} \Comment{Worst case loop over every vertex}
            	 \For{every vertex $v \in G(V)$}
            	 	\State calculate neighborhood $nbs$ of $v$
            	  	\If {$nbs$ of $v$ has not changed}
    \State pass
    				\Else
    					\For{every item in the dictionary}
    						\State checks if the configuration is not yet known.
    					\EndFor
    					\If{the configuration is not yet known}
    						\State assign a new color \
    					\EndIf
    				\EndIf
            	  \EndFor
				\For{every vertex in G}
					\State update neighborhood dictionary
				\EndFor
            \EndWhile\label{euclidendwhile}
            \State \textbf{return} stable colored graph
        \EndProcedure
    \end{algorithmic}
    \caption{Basic color refinement}
    \label{euclid}
\end{algorithm}

%------------------------------------------------
\section{Figures}
\label{app:figs}
\begin{multicols}{2}
\begin{figure}[H]
	\includegraphics[width=\linewidth]{basiconly}
	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
	\label{fig:basiconly}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{fastonly}
	\caption{The computation times of the fast color refinement algorithm for various sizes of the three paths graph.}
	\label{fig:fastonly}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{basicfunctions}
	\caption{The computation times of the basic color refinement algorithm and the functions $f(n)=n^2$ and $f(n)=n^3$.}
	\label{fig:basicfunctions}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{basicdiff}
	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(n)=n^2$ and $f(n)=n^3$.}
	\label{fig:basicgrowth}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{fastfunctions}
	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
	\label{fig:fastfunctions}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{fastdiff}
	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
	\label{fig:fastgrowth}
\end{figure}
\end{multicols}

%------------------------------------------------
% \newpage
\section{Tables} %Change section name
\label{app:tabs}
\begin{table}[h]\centering
\begin{tabular}{|l|r@{.}l|r@{.}l|}
\hline
Graph	
	& \multicolumn{2}{l|}{Basic \emph{(s)}}
	& \multicolumn{2}{l|}{Fast \emph{(s)}} \\
\hline
threepaths\_{}5
	& 0 & 00045
	& 0	& 00059 \\
threepaths\_{}10
	& 0 & 0017
	& 0	& 00086 \\
threepaths\_{}20
	& 0 & 0072
	& 0	& 0018 \\
threepaths\_{}40
	& 0 & 031
	& 0	& 0044 \\
threepaths\_{}80
	& 0 & 126
	& 0	& 0071 \\
threepaths\_{}160
	& 0 & 604
	& 0	& 014 \\
threepaths\_{}320
	& 3 & 213
	& 0	& 029 \\
threepaths\_{}640
	& 19 & 261
	& 0	& 058 \\
threepaths\_{}1280
	& 127 & 567
	& 0	& 117 \\
threepaths\_{}2560
	& 907 & 407
	& 0	& 261 \\
threepaths\_{}5120
	& 5736 & 550
	& 0	& 491 \\
threepaths\_{}10240
	& \multicolumn{2}{c|}{-}
	& 0	& 994 \\
threepaths\_{}10240
	& \multicolumn{2}{c|}{-}
	& 2	& 089 \\

\hline
\end{tabular}
\caption{The measured computation times of the basic and fast color refinement algorithms for various sizes of the three paths graph.}
\label{table:measurements}
\end{table}
%------------------------------------------------

\end{appendices}
%------------------------------------------------

\end{document}